\chapter{Conclusion}

\section{Contributions}

This diploma project provides the following contributions to the domain of computer science:
\begin{itemize}
	\item An indefinitely extensible image filter library, that can be used alongside any user interface.
	\item A graphical interface suitable for tweaking each provided parameter to the need of the user.
	\item A measurement utility for computing filter delay and performing system diagnostics and comparisons.
	\item An application that makes use of the aforementioned libraries in order to provide an insightful and
	      educational introduction to the field of image signal processing.
\end{itemize}

\section{Impact}

The main result of this project is and educational application that can be used as tools in the following
subdomains:
\begin{itemize}
	\item Machine learning - with the convolutional pooling methods being exemplified by the box blur, median
	      blur, erode and dilate.
	\item 2D frequency domain analysis - with the Gaussian and bilateral smoothing techniques providing an
	      introduction to the field.
	\item Computer Vision - with the Sobel and Canny edge detectors acting as an intermediary between the
	      aforementioned subdomains.
	\item Industry application - with the emboss filter allowing for the preview of embossed metal stamps
	      before they are produced.
	\item Optics and photography - with the lens filter being both a useful tool for camera calibration, as
	      well as a creative mean for introducing intentional distortion in images.
\end{itemize}



